Master Thesis

### 2013-06-13 the-last-part.tex
More formal talking about GACGs

\section{Putting It All Together}

@@

    \item Let $L_1$ and $L_2$ be two extrinsic languages defined by the
      nodes $u_1$ and $u_2$ respectively belonging to two graphical
      ACGs. We can construct a new graphical ACG by taking the union of
      the two ACGs (union of graphs (union of vertices and edges) and
      unions of the labelings).\footnote{Assuming, without loss of
        generality, that the sets of vertices of the two graphs are
        disjoint.} We then add a new node $v$ which will define the
      language $L_1 \cap L_2$.

@@

    \begin{align*}
      \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v) \cap
      \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
      &= \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
      &= \mathcal{L}(\mathcal{E}_{\mathcal{G}}(u)) \\
      &= \mathcal{L}(L)
    \end{align*}
  \end{itemize}
\end{proof}

@@

>>>>

\section{Putting It All Together}
\label{sec:together}

@@

    \item Let $L_1$ and $L_2$ be two extrinsic languages defined by the
      nodes $u_1$ and $u_2$, respectively, and belonging to some two
      graphical ACGs. We can construct a new graphical ACG by taking the
      union of the two ACGs (union of graphs (union of vertices and
      edges) and unions of the labelings).\footnote{Assuming, without
        loss of generality, that the sets of vertices of the two graphs
        are disjoint.} We then add a new node $v$ which will define the
      language $L_1 \cap L_2$.

@@

    \begin{align*}
      \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v) \cap
      \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
      &= \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
      &= \mathcal{L}(\mathcal{E}_{\mathcal{G}}(u)) \\
      &= \mathcal{L}(L)
    \end{align*}

    The facts used here are in turn: the inclusion of
    $\mathcal{I}_{\mathcal{G}'}(v)$ in
    $\mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u))$, the observation that
    modifying our graphical grammar did not change the extrinsic
    language of $u$ and the assumption that $L$ is the extrinsic
    language of $u$.
  \end{itemize}
\end{proof}

@@

We conclude this subsubsection with a few words to motivate the need we
feel for this (possibly) more expressive formalism. We believe that in
order to manage the development of wide-coverage grammars, the grammars
need to be kept as simple as possible, which means, among other things,
that independent constraints should be implemented separately and not
complected together (we saw the difficulties caused by this already on a
microscopic fragment in section~\ref{sec:together}). This means that if
we were to define, e.g., a partial grammar in which we would only
enforce agreement and another one in which we would only enforce the
proper use of relative pronouns, we would like the language in which
both agreement and proper use of relative pronouns are respected to be
expressible in our formalism and on top of that, we would like this
language to be easily expressed in terms of the two grammars for the
individual constraints. We now have a requirement on the formalism in
which we would write such grammars, we want a formalism where the set of
all the ``useful'' definable languages is closed on intersection. If the
object languages of ACGs are indeed closed on intersection, then our
extension is just a convenience for expressing intersection more
easily. However, if they are not, we believe our extension adds a
critical level of expressivity needed for simple and modular
wide-coverage grammars.

\subsubsection{On the Decidability of Graphical ACGs}

We turn to the problem of parsing with graphical ACGs. Parsing with ACGs
is a problem on the limits of decidability. When we consider ACGs whose
abstract signatures only contain types of order 2 or less, parsing is
possible in polynomial type for both linear \cite{salvati2005problemes}
and non-linear \cite{kanazawa2007parsing} type systems (such as the one
we introduced in~\ref{sec:acg}). If the order of the abstract signature
exceeds 2, the membership problem for linear ACGs becomes decidable if
and only if the satisfiability problem in the multiplicative exponential
fragment linear logic is decidable \cite{de2004vector}, which is an open
problem. On the other hand, by extending the type system to include
dependent types, parsing can become undecidable too since type checking
in such a system is itself undecidable.

Our intent here is to show that by extending the formalism, we have not
made the problem of parsing with the grammar more difficult, at least in
terms of decidability if not in terms of complexity. Let us imagine a
graphical ACG where every edge (two signatures, distinguished types and
a lexicon) is an ACG for which there is an effective parsing algorithm
and all type checking is decidable in all the signatures. Then we can
build a simple effective parsing algorithm for any of the extrinsic
languages defined in the graphical ACG.

We will be testing whether the term $\gamma$ belongs to the extrinsic
language of some node $v$. Our algorithm first checks whether $\gamma$
belongs to the intrinsic language of $v$ by checking its type. If it
does, it proceeds to verify that it is also in the extrinsic language by
finding antecedents in all of its abstract predecessors $u_i$, if there
are any. We can use the presupposed parsing algorithm for the ACG
$\mathopen{<} \Sigma_{u_i}, \Sigma_v, \mathcal{L}_{(u_i,v)},
S_{u_i}\mathclose{>}$ to find all the abstract terms $\alpha_{i,j}$
having type $S_{u_i}$ and being antecedents to $\gamma$ w.r.t. the
lexicon $\mathcal{L}_{(u_i,v)}$. Such terms belong to the intrinsic
language of $u_i$ but we are interested only in antecedents coming from
the extrinsic language. Therefore, what we have to do is to recursively
apply our parsing algorithm to all of the antecedent-candidates
$\{\alpha_{i,j} \mid \forall j\}$ until we find one which belongs to the
extrinsic language of $u_i$. Since the graph is acyclic, we are
guaranteed that this recursive traversal will terminate. We repeat this
for all the predecessors $u_i$ and if we succeed to find antecedents in
all of them, we have confirmed the membership of $\gamma$ in
$\mathcal{E}_{\mathcal{G}}(v)$ and also built its abstract
antecedent(s).

\subsubsection{On Grammar Engineering with Graphical ACGs}

%% \subsubsection{On Alternative Interpretations of Graphical ACGs}
%
% Mention the example with the levels of linguistic representations
% linked by many-to-many transductions and explain how a different
% theory of graphs defining languages might be suitable to that
% architecture.
%
% Also point out the weirdness of the diamond shape.
